### 2011-07-11 Paper/Paper.tex
Updated paper. Added support for never occuring observations although not sure if it is working. Also: logistic discretisation! And more relative features

\begin{abstract}
Both machine learning and rule based techniques have been extensively applied to performance rendering. However, relatively few systems make explicit use of musical structure. This paper introduces a performance rendering approach that tries to learn expression exclusively at a structural level. The approach attempts to to be complementary to approaches that learn expression at note level. Musical structure is extracted automatically. A dataset containing performances aligned to scores is then used to learn how structure relates to expression. Finally, the model can be applied to a a new score to produce an expressive performance of the score.

@@

It is customary in Western art music tradition that composers write down their compositions in scores. The task of a performer is to some extend to accurately reproduce the score, however, a perfect reproduction of a score generally sounds robotic and unpleasant. What makes a performance appealing is that the performer deviates from the score, altering timing, articulation and loudness, creating an expressive performance.

@@

The YQX system, as widmer admitted, tended to sometimes produce nervous sounding changes in expression. He presents two extensions that seek to generate smoother performances. The problem with these extensions is, as widmer admits himself, that the increased smoothness comes at the expense of expressitivity. To counter this, he adds three explicit rules that he uses to postprocess the performances. We think the reason that Widmer stumbled upon this tradeoff between expressiveness and nervousness because the nervousness is inherent to the way performances are generated at note-level. The system simply misses one component of expression, namely structure, that Widmer had to artificially introduce using rules.

@@

We can now convert a piece of music into a serie of constituents. These constituents will be used to predict expression so we must be able to charaterize them in a way correlates with the way they are performed. Clearly, expressive markings in the score play a role in how the constituent should be played. If the segmentation is specific enough  there will hopefully be at most expressive marking within the constituent.

\section{Performance Rendering}
\section{Dealing with polyphony}
\subsection{Dataset and representation}

@@

\section{Future Work}
\subsection{Repetition}
A notion of repetition and similarity would certainly improve the system. Repetition is a very good indicator of constituent breaks. Repetition and similarity could also be used to improve expressiveness of performances. It is probably telling when a phrase is repeated three times and then slightly altered the fourth time. Although finding similarity and musically significant repetition is a subject of its own the delta framework could help to define repetition arbitrarily of transposition or rhythm. A list of pitch deltas can for example be used to detect repetition independend of transposition and a list of duration deltas can be used to detect repetition of rhythm independend of the notes used.

>>>>

\begin{abstract}
Both machine learning and rule based techniques have been extensively applied to performance rendering. However, relatively few systems make explicit use of musical structure, combined with machine learning. This paper introduces a performance rendering approach that tries to learn expression exclusively at a structural level. The approach attempts to to be complementary to approaches that learn expression at note level. Musical structure is extracted automatically. A dataset containing performances aligned to scores is then used to learn how structure relates to expression. Finally, the model can be applied to a a new score to produce an expressive performance of the score.

@@

In Western art music tradition it is customary that composers write down their compositions in scores. The task of a performer is to some extend to accurately reproduce the score, however, a perfect reproduction of a score generally sounds robotic and unpleasant. What makes a performance appealing is that the performer deviates from the score, altering timing, articulation and loudness, creating an expressive performance.

Given a hypothetical perfect synthesizer, performing music with computers is a trivial task. However \textit{expressively} performing music is not and much research has focussed on this issue. Computers are widely used in music production. Since support for automatic expression is mostly absent or very poor in music editing software, some genres of music have evolved to sound acceptable even without expression. If automatic expression better in this software, computer music production could greatly widen its field of application.

Some music transcription
Performance rendering

@@

The YQX system, as widmer admitted, tended to sometimes produce nervous sounding changes in expression. He presents two extensions that seek to generate smoother performances. The problem with these extensions is, as widmer admits himself, that the increased smoothness comes at the expense of expressitivity. To counter this, he adds three explicit rules that he uses to postprocess the performances. We think the reason that Widmer stumbled upon this tradeoff between expressiveness and nervousness because the nervousness is inherent to the way performances are generated at note-level. The system simply misses one component of expression, namely structure level expression. (that Widmer had to artificially introduce using rules. ) ?

@@

We can now convert a piece of music into a serie of constituents. These constituents will be used to predict expression so we must be able to charaterize them in a way correlates with the way they are performed. Analogous to YQX we are looking for the \textit{context} of the constituent.

(Clearly, expressive markings in the score play a role in how the constituent should be played. If the segmentation is specific enough  there will hopefully be at most expressive marking within each constituent. )?

\section{Targets}

\begin{itemize}
\item Use average expression parameters per constituent
\end{itemize}

\section{Performance Rendering}
\begin{itemize}
\item Extract melody
\item Extract scorefeatures from melody
\item lookup notes in deviation, ignore missing notes
\item Extract expressive and non-expressive melody and extract expression features
\item Perform notes other than melody notes the same as the nearest melody note
\end{itemize}
\subsection{Dataset and representation}
\begin{itemize}
\item attack release
\item tempo curves
\end{itemize}
\subsection{Dealing with polyphony}
\subsection{Discretization}

@@

\begin{itemize}
\item Top notes is not always the melody, musical attention
\item Bass and harmony shouldn't be played with the same expression as melodynotes
\end{itemize}

\section{Future Work}
\begin{itemize}
\item Use loudness and tempo direction instead of averages (requires more data)
\item Generalize approach to incorporate hierarchical structure, let structure extend to note level, so note level expression and structurelevel expression become integrated
\end{itemize}
\subsection{Repetition}
A notion of repetition and similarity would certainly improve the system. Repetition is a very good indicator of constituent breaks. Repetition and similarity could also be used to improve expressiveness of performances. It is probably telling when a phrase is repeated three times and then slightly altered the fourth time. Although finding similarity and musically significant repetition is a subject of its own the delta framework could help to define repetition arbitrarily of transposition or rhythm. A list of pitch deltas can for example be used to detect repetition independend of transposition and a list of duration deltas can be used to detect repetition of rhythm independend of the notes used.
